---
title: "data cleaning"
output: pdf_document
---

merging the files
```{r}
install.packages("dplyr")
library(dplyr)
install.packages("data.table")
library(data.table)

fixations = fread("FixationsV1.csv")
saccades = fread("SaccadesV1.csv")
samples = fread("SamplesV1.csv")
log1 = fread("logfile_1_2_f.csv")
log2 = fread("logfile_2_1_f.csv")
log3 = fread("logfile_3_2_f.csv")
log4 = fread("logfile_4_1_F.csv")
log5 = fread("logfile_5_2_m.csv")
log6 = fread("logfile_6_1_m.csv")

logfiles = rbind(log1, log2, log3, log4, log5, log6)

logfiles$V1 = as.factor(logfiles$V1)
logfiles$V1 = as.numeric(logfiles$V1)

logfiles= rename(logfiles, "Trial" = "V1")

logfiles=rename(logfiles, "ParticipantID" = "subject")

#substr gender, ostensiveness, direction

logfiles$actorGender = substr(logfiles$video, start = 1, stop = 1)
logfiles$ostensiveness = substr(logfiles$video, start = 13, stop = 14)
logfiles$direction = substr(logfiles$video, start = 9, stop =  11)


## merging logfiles with V1's

fixationsM = merge(fixations, logfiles, by = c("ParticipantID", "Trial"), all = T)
saccadesM = merge(saccades, logfiles, by = c("ParticipantID", "Trial"), all = T)
samplesM = merge(samples, logfiles, by = c("ParticipantID", "Trial"), all = T)

# adding condition
fixationsM$condition = NA
samplesM$condition = NA
saccadesM$condition = NA

samplesM$condition[which(samplesM$SearchOrder == "1" & samplesM$Trial <6)]= "Star"
samplesM$condition[which(samplesM$SearchOrder == "2" & samplesM$Trial >5)]= "Count"
samplesM$condition[which(samplesM$SearchOrder == "2" & samplesM$Trial <6)]= "Star"
samplesM$condition[which(samplesM$SearchOrder == "1" & samplesM$Trial >5)]= "Count"

fixationsM$condition[which(fixationsM$SearchOrder == "1" & fixationsM$Trial <6)]= "Star"
fixationsM$condition[which(fixationsM$SearchOrder == "2" & fixationsM$Trial >5)]= "Count"
fixationsM$condition[which(fixationsM$SearchOrder == "2" & fixationsM$Trial <6)]= "Star"
fixationsM$condition[which(fixationsM$SearchOrder == "1" & fixationsM$Trial >5)]= "Count"

saccadesM$condition[which(saccadesM$SearchOrder == "1" & saccadesM$Trial <6)]= "Star"
saccadesM$condition[which(saccadesM$SearchOrder == "2" & saccadesM$Trial >5)]= "Count"
saccadesM$condition[which(saccadesM$SearchOrder == "2" & saccadesM$Trial <6)]= "Star"
saccadesM$condition[which(saccadesM$SearchOrder == "1" & saccadesM$Trial >5)]= "Count"


```

modelling the data:
Searchtype is controlling for the different order the pictures/conditions are shown
Systemtic change over time per trial is controlled for by adding trial as a 
Adding fixations controls for the temporal effects in eack trial - each fixation is a datapoint, i.e. durations of these may differ systematically over time.

log transformation:
family=gaussian(link=log)


```{r}
# model w. everything predicting fixation duration
# F_duration ~ SearchType * SearchOrder * fixations * Trial + (1+ST*fix*Trial| ParticipantID)

#simpel model
#F_duration ~ searchType * Trial + (1+ST*Trial|PID)

library(Metrics)
library(caret)
library(lme4)
library(lmerTest)
library(modelr)

# subsetting
data = subset(fixationsM, Task == "VisualSearch")

data$ParticipantID = as.character(data$ParticipantID)
data$ParticipantID = as.factor(data$ParticipantID)
data$ParticipantID = as.numeric(data$ParticipantID)

data$folds_id = as.numeric(data$ParticipantID)

flds <- createFolds(unique(data$ParticipantID), k = 3)

test_rmse = NULL
train_rmse = NULL
n =  1

for(i in flds){
  
  #test
  testsub = subset(data, folds_id %in% i)
  #train
  trainsub = subset(data,!(folds_id %in% i))
  #train model
  trainm = lmer(Duration ~ 1 + condition + (1+condition + Trial|ParticipantID), data = trainsub, REML = F)
  
  # predicting values for the dependent variable from the testdata using the training model 
  pred = predict(trainm, testsub, allow.new.levels = T)
  
  # calculating the rmse between the actual values in CHI_MLU and the predicted values for both the test and the training subset  
  train_rmse[n]= Metrics::rmse(trainsub$Duration, fitted(trainm)) 
  test_rmse[n]=Metrics::rmse(testsub$Duration, pred)
  
  
    
  n=n+1
}

mean(test_rmse)
mean(train_rmse)
```

#predicting pupil size:
 pupilSize ~ ost + dir
 
fixation dataset - pupil size is averaged over the individual fixations
samples dataset - contains information about pupil size for each ms, i.e. a lot of data!(a solution could be downsampling the data - a pragmatical good idea, but as the pupil size changes on a scale of ms. the information would be more complete if avoided)

## Using the sample data:

  PupilSize ~ 1 + ostensiveness * direction + actorGender * ParticipantGender + Trial + TimeTrial + (1+ Trial + TimeTrial + Ost + direction |ParticipantID)
  
  because of the luminance of the video the pupil size will be affected at the beginning and the end of each trial, which can be tested for using the TimeTrial
  direction and ostensiveness may affect each other - 
  
  














